---
title: "SpokaneProviders"
author: "Issy Belzil"
date: '2017-10-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
head(providerspokane)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
load(providerspokane)
```

### 1.Are there differences in the number of services, distinct beneficiary per day services, average medicare allowed, charged, and paid amount, and medicare standardized amount differ as a function of Gender, the Provider Type, and Place of Service.


#### Studying Gender as a Function of: Number of services, distinct beneficiary per day services, average medicare allowed, charged and paid amount and medicare standardized amount 

Variables: () 

-Number of Services 1 
-Distinct beneficiary per day services 1
-Average medicare allowed 1
-Charged 1 
-Paid amount 1 
-Medicare standardized amount 1 


```{r}
Genderstudy=providerspokane %>% select( Gender.of.the.Provider, Number.of.Distinct.Medicare.Beneficiary.Per.Day.Services, Average.Medicare.Allowed.Amount, Average.Submitted.Charge.Amount, Number.of.Services, Average.Medicare.Standardized.Amount, Average.Medicare.Payment.Amount)
```

```{r}
str(Genderstudy)
```

```{r}
ggplot(Genderstudy, aes(x=reorder(Gender.of.the.Provider, Number.of.Distinct.Medicare.Beneficiary.Per.Day.Services),  y=Number.of.Distinct.Medicare.Beneficiary.Per.Day.Services, )) + geom_bar(stat="identity", alpha=.60)+coord_flip() + labs(list(title = "Gender Study", x = "Gender of Provider", y = "Number of Distinct Medicare Ben. Per Day Service"))
```

Take out NA from Gender 

Try: 
facet grids 


#### Studying the Provider Type as a Function of: Number of services, distinct beneficiary per day services, average medicare allowed, chargedm and paid amount and medicare standardized amount 

Provider.Type


```{r}
Providerstudy=providerspokane %>% select( Provider.Type, Number.of.Distinct.Medicare.Beneficiary.Per.Day.Services, Average.Medicare.Allowed.Amount, Average.Submitted.Charge.Amount, Number.of.Services, Average.Medicare.Standardized.Amount, Average.Medicare.Payment.Amount)
```

#### Studying the Place of Service as a Function of: Number of services, distinct beneficiary per day services, average medicare allowed, chargedm and paid amount and medicare standardized amount 

Place.of.Service

```{r}
PlaceofServicestudy=providerspokane %>% select( Place.of.Service, Number.of.Distinct.Medicare.Beneficiary.Per.Day.Services, Average.Medicare.Allowed.Amount, Average.Submitted.Charge.Amount, Number.of.Services, Average.Medicare.Standardized.Amount, Average.Medicare.Payment.Amount)
```



# 2.How do all these variables vary as a function of the type of service that is being provided?


